---
title: "FinalProjectBiology4415"
author: "Gavin Ieong"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r loading librarys and files, echo=FALSE, message= F}
library(readxl)
library(ggplot2)
library(reshape2)
library('corrr')
library(ggcorrplot)
library("FactoMineR")
library("factoextra")
library(ggfortify)
library(RColorBrewer)
library(dplyr)
#library(VennDiagram)
#library(ggvenn)
library(pheatmap)
library(ggpubr)
```



## Part I: Different Visualization Techniques
We are doing this to make it easier to see changes compared to the stackerd bar graph in original paper. Helps aleviate the unlevelness and with a better colour scheme



```{r bubbleplot protein, warning=F}
table10 <- as.data.frame(read_excel("table.XLSX", "Supplemental Table 10"))
colnames(table10) <- as.character(unlist(table10[1,]))
table10 <- table10[-1, ]

colnames(table10)[1] <- "Compartment"

bubbleProtein <- table10[, c(1,3,5,9,13,7,11,15)]

bubbleProtein[,(2:8) ] <- sapply(bubbleProtein[,(2:8) ] , as.numeric)
colnames(bubbleProtein) <- c("Compartment","Total","Line14 Up","Line15 Up","Line16 Up","Line14 Down","Line15 Down","Line16 Down")

bubbleProtein$Compartment <- factor(bubbleProtein$Compartment,levels=unique(bubbleProtein$Compartment))

bubbleProtein <- bubbleProtein[1:10,]
pcm = melt(bubbleProtein, id = c("Compartment"))
colours = c( "#A54657",  "#582630", "#F7EE7F", "#4DAA57",
           ## "#F1A66A","#F26157", "#F9ECCC", "#679289", "#33658A",
"#F6AE2D","#86BBD8","#B9AC2A") ##,"#C6BBD8""#8DFE2D")




ggplot(pcm, aes(x = Compartment, y = variable)) + 
  geom_point(aes(size = value, fill = variable), alpha = 0.75, shape = 21) + 
  scale_size_continuous(limits = c(0, 50), range = c(0,15), breaks = c(0,10, 20,30,40,50)) + 
  labs( x= "Location Of Protein", y = "Regulation of Transgenic Line", title = "Localization of Proteins", size = "Percentage Of Proteins", fill = "")  + 
  theme(legend.key=element_blank(), 
  axis.text.x = element_text(colour = "black", size = 12, face = "bold", angle = 90, vjust = 0.3, hjust = 1), 
  axis.text.y = element_text(colour = "black", face = "bold", size = 11), 
  legend.text = element_text(size = 10, face ="bold", colour ="black"), 
  legend.title = element_text(size = 12, face = "bold"), 
  panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = .5), 
  legend.position = "right") +  
  scale_fill_manual(values = colours, guide = FALSE) + 
  scale_y_discrete(limits = rev(levels(pcm$variable)))

##ggsave("Protein_Bubble_Plot.png")
```



## Pca plot of protein intensities in each of the three transgenic lines, plus wild type replicates
Will produce PCA plot first, to confirm we have similar data compared to original paper
```{r PCA analysis}
table9 <- as.data.frame(read_excel("table.XLSX", "Supplemental Table 9"))
colnames(table9) <- as.character(unlist(table9[1,]))
table9 <- table9[-1, ]

table9[5:22] <- sapply(table9[5:22] , as.numeric)
allTrials <- table9[c(5:16)]
allTrials[allTrials==0] <- NA
allTrials<-allTrials[complete.cases(allTrials),]

data_normalized <- scale(allTrials)
corr_matrix <- cor(data_normalized)

## Can change col names after msking corr matrix to use as groups
colnames(allTrials)<- c("Wildtype","Wildtype","Wildtype","Line14","Line14","Line14","Line15","Line15","Line15","Line16","Line16","Line16")
data.pca <- princomp(corr_matrix)

group.colors <- c(Wildtype = "purple", Line14 = "blue", Line15 ="red", Line16 = "green")


fviz_pca_ind(data.pca,repel = T , habillage=colnames(allTrials)) +
  # ADD ggforce's ellipses
  ggforce::geom_mark_ellipse(aes(fill = Groups))  + theme_minimal()+scale_color_manual(values=group.colors)+scale_fill_manual(values=group.colors)

```

## Part II: Plot some basic information before looking at other graphs
Raw scatterplot of intensities compared to wildtype. Used as baseline, but will further use other plots to help confirm and find information.
```{r}


intensity <- as.data.frame(read_excel("table.XLSX", "AverageIntensity"))
intensity<- intensity[!duplicated(intensity$AGI), ]
rownames(intensity) <- intensity[,1]
intensity <- intensity[,-1]
intensity<- log10(intensity)

names(intensity)
#intensity<-head(intensity,100)
par(mfrow=c(1,3))

ggscatter(intensity, x =  "Average Intensity Of Wildtype Line", y =  "Average Intensity Of  Line 14" ,
   add = "reg.line",  # Add regressin line
      color = "blue",
   add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )+ stat_cor(
   aes(label = paste(..rr.label.., sep = "~`,`~")),
  label.x = 3
)

ggscatter(intensity, x =  "Average Intensity Of Wildtype Line", y =  "Average Intensity Of  Line 15" ,
   add = "reg.line",  # Add regressin line
   color = "red",
   add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )+ stat_cor(
   aes(label = paste(..rr.label.., sep = "~`,`~")),
  label.x = 3
)

ggscatter(intensity, x =  "Average Intensity Of Wildtype Line", y =  "Average Intensity Of Line 16" ,
   add = "reg.line",  # Add regressin line
      color = "green",
   add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )+ stat_cor(
   aes(label = paste(..rr.label.., sep = "~`,`~")),
  label.x = 3
)

plot(intensity[c(1,2)], col = "blue",pch = 8, main = "Avg Intensity of Line 14 vs Wildtype")

plot(intensity[c(1,3)], col = "red",pch = 19, main = "Avg Intensity of Line 15 vs Wildtype")

plot(intensity[c(1,4)], col = "#53C436",pch = 10, main = "Avg Intensity of Line 16 vs Wildtype")

```


## Visualize Foldchange of each lines using histogram
Distribution of log2 Fold Change in lines 14,15, 16
Line 15 and Line 16 seem to match more compared to Line 14, where more differences are noticed
```{r}

topProteins <- as.data.frame(read_excel("table.XLSX", "TopGenes"))
par(mfrow=c(1,3))
hist(topProteins$FC_14_WT, xlab = "Log2 Fold Change (Line14 vs Wildtype)", main = "Protein Log 2 Foldchange Line 14",col = 'blue', border = "white")
hist(topProteins$FC_15_WT, xlab = "Log2 Fold Change (Line15 vs Wildtype)", main = "Protein Log 2 Foldchange Line 15",col = 'red', border = "white")
hist(topProteins$FC_16_WT, xlab = "Log2 Fold Change (Line16 vs Wildtype)", main = "Protein Log 2 Foldchange Line 16",col = 'green', border = "white")

```


## Heat map Of Log2Fold Change Against Wildtype 
 Once again, as seen from heatmap, line 14 shows most significant difference
```{r heatmap}
myheatmap<- table9[c(4,20,21,22)]
myheatmap<-myheatmap[!duplicated(myheatmap$AGI),]
myheatmap <-na.omit(myheatmap)
pheatmap(as.matrix((myheatmap[2:4])), scale="row", show_rownames=F,legend = TRUE, main = "Log2 Fold Change Value of Each Protein Per Transgenic Line")

```



#### Bubble Plot (Gene Count) This section removed, as no need to do genes when working with proteomics.
```{r loading Gene bubblePlot, eval = F}
table1 <- as.data.frame(read_excel("table.XLSX", "Supplemental Table 1"))
colnames(table1) <- as.character(unlist(table1[1,]))
table1 <- table1[-1, ]
#rownames(table1)<- table1[,1]
#table1<- table1[,-1]
table1[,(2:15) ] <- sapply(table1[,(2:15) ] , as.numeric)

bubblePlotGene <- table1[c(1,3,5,7,9,11,13,15)]
colnames(bubblePlotGene) <- c("Compartment","Total","Line14 Up","Line15 Up","Line16 Up","Line14 Down","Line15 Down","Line16 Down")

pcm = melt(bubblePlotGene, id = c("Compartment"))

colours = c( "#A54657",  "#582630", "#F7EE7F", "#4DAA57",
           ## "#F1A66A","#F26157", "#F9ECCC", "#679289", "#33658A",
"#F6AE2D","#86BBD8","#B9AC2A") ##,"#C6BBD8""#8DFE2D")

bubblePlotGene$Compartment <- factor(bubblePlotGene$Compartment,levels=unique(table$bubblePlotGene))


ggplot(pcm, aes(x = Compartment, y = variable)) + 
  geom_point(aes(size = value, fill = variable), alpha = 0.75, shape = 21) + 
  scale_size_continuous(limits = c(0, 100), range = c(0,10), breaks = c(0,20, 40, 60, 80, 100)) + 
  labs( x= "Location Of Gene", y = "Regulation of Transgenic Line", title = "Localization of Genes", size = "Gene Count Percentage", fill = "")  + 
  theme(legend.key=element_blank(), 
  axis.text.x = element_text(colour = "black", size = 12, face = "bold", angle = 90, vjust = 0.3, hjust = 1), 
  axis.text.y = element_text(colour = "black", face = "bold", size = 11), 
  legend.text = element_text(size = 10, face ="bold", colour ="black"), 
  legend.title = element_text(size = 12, face = "bold"), 
  panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = .5), 
  legend.position = "right") +  
  scale_fill_manual(values = colours, guide = FALSE) + 
  scale_y_discrete(limits = rev(levels(pcm$variable)))
  

```


#### Citations
Used to get citations
```{r }
citation('pheatmap')
citation('ggplot2')
citation('corrr')
citation('ggcorrplot')
citation('FactoMineR')
citation('factoextra')
citation('ggfortify')
citation('ggpubr')
```